---
title: "Homework 1"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:** 1-5 (3.2.4 Exercises); 1-2, 5 (3.3.1 Exercises); 1-5 (3.6.1 Exercises); Open Response

**Submission:** Submit an electronic document on Gradescope. Must be submitted as a PDF file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
```

# Chapter 3 (3.2.4 Exercises)

##  Exercise 1
```{r}
#
```


## Exercise 2
```{r}
#
```


##  Exercise 3
```{r, eval=FALSE}
?mpg
```

```{r}
#
```


##  Excercise 4
```{r}
#
```


##  Excercise 5
```{r}
#
```


# Chapter 3 (3.3.1 Exercises)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

##  Exercise 5
```{r}
#
```

# Chapter 3 (3.6.1 Exercises)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 4
```{r}
#
```

##  Exercise 5

I don't know if they will look different. Let me check.

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```

They do not look different. I am incredibly surprised.

# Open Response

For this exercise, use the **diamonds** dataset in the tidyverse. Use `?diamonds` to get more information about the dataset.

## Step 1: Select 1 numeric variable and 2 categorical variables. Create a graphic using `geom_boxplot()` and `facet_wrap` to illustrate the empirical distributions of the sample.


## Step 2: Choose 2 numeric variables and 2 categorical variables and creatively illustrate the relationship between all the variables. 
