---
title: "Homework 3"
author: "FIRSTNAME LASTNAME"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Instructions

**Exercises:**  1,3 (7.3.4 Exercises); 1 (7.4.1 Exercises); 2,4 (7.5.1.1 Exercises); 1,2 (7.5.2.1 Exercises); 2,3,5 (7.5.3.1 Exercises)

**Submission:** Submit an electronic document on Gradescope. Must be submitted as a PDF file generated in RStudio. All assigned problems are chosen according to the textbook *R for Data Science*. You do not need R code to answer every question. If you answer without using R code, delete the code chunk. If the question requires R code, make sure you display R code. If the question requires a figure, make sure you display a figure. A lot of the questions can be answered in written response, but require R code and/or figures for understanding and explaining.

```{r, include=FALSE}
library(tidyverse)
```

# Chapter 7 (7.3.4 Exercises)

##  Exercise 1
```{r}
#
```

##  Exercise 3
```{r}
#
```

# Chapter 7 (7.4.1 Exercises)

##  Exercise 1
```{r}
#
```

# Chapter 7 (7.5.1.1 Exercises)

##  Exercise 2
```{r}
#
```

##  Exercise 4
```{r}
#
```

# Chapter 7 (7.5.2.1 Exercises)

##  Exercise 1
```{r}
#
```

##  Exercise 2
```{r}
#
```

# Chapter 7 (7.5.3.1 Exercises)

##  Exercise 2
```{r}
#
```

##  Exercise 3
```{r}
#
```

##  Exercise 5
```{r}
#
```
